Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-02-03 Multiresolution_framework/ltx/manuscript.tex
corrected the appendix format

The multi-resolution approximation \citep{Mallat1989a} of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(\cdot)$, and the connectivity kernel, $w(\cdot)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(\mathbf{r})$ and a mother wavelet $\psi(\mathbf{r})$. Considering a  one-dimensional field, the static connectivity kernel is decomposed as,
\begin{equation}
 w\left(r\right)=\sum_{l\in \mathbb{Z}}\alpha_{j_0,l} \underbrace{\phi_{j_0,l}\left(r\right)}_{\text{scaling functions}} + \sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}}\beta_{j,l} \underbrace{\psi_{j,l}\left(r\right)}_{\text{wavelets}}, 
\label{eq:KernelExpansion}
\end{equation}
where $\alpha_{j_0,l}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,l}$ are the detail coefficients at different scales $j$, with
\begin{align}
\phi_{j,l}\left(r\right) =& 2^{\frac{j}{2}}\phi\left(2^j r-l\right) \\	
\psi_{j,l}\left(r\right) =& 2^{\frac{j}{2}}\psi\left(2^j r-l\right).
\end{align}
The parameters $j$ and $l$ control the scale (dilation) and translation (spatial shift), respectively. \dean{NEED SOMETHING ABOUT HOW j AND l ARE SET HERE - AS IN THAT THEY ARE KNOWN - AND THAT WE ONLY NEED TO ESTIMATE THE COEFFICIENTS.} \parham{The number of spatial shifts is dependent on the scale, $j$, and can be chosen such that the basis functions cover the spatial domain of interest. The number of scales depends on the spectral characteristics of the field, which will be described later.} Scaling functions retain the lowest frequency components, due to their low-pass properties, while the wavelet functions extract successively higher frequency components due to their band-pass characteristics.

% At each spatial scale indexed by $j$ (level of decomposition), the neural field can be written as
% \begin{equation}
% 	v_{j,t}\left(\mathbf{r}\right) = \begin{cases}
% 	\sum_{l \in \mathbb{Z}}x_{j,l}(t) \underbrace{\phi_{j,l}\left(\mathbf{r}\right)}_{\text{scaling function}} + \check{x}_{t,j,l}\underbrace{\psi_{j,l}\left(\mathbf{r}\right)}_{\text{mother wavelet}} & \text{if $j=0$},\\
% 	\sum_{l \in \mathbb{Z}}\check{x}_{t,j,l}\psi_{j,l}\left(\mathbf{r}\right) & \text{otherwise}
% \end{cases}
% \end{equation}

The dynamic neural field is decomposed using
\begin{equation}
 v_t\left(r\right)=\sum_{l \in \mathbb{Z}}x_{t,j_{0},l} \underbrace{\phi_{j_0,l}\left(r\right)}_{\text{scaling functions}} + \sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}^n} \check{x}_{t,j,l} \underbrace{\psi_{j,l}\left(r\right)}_{\text{wavelets}},
\label{eq:FieldExpansion}
\end{equation}
where $x_{t,j_{0},l}$ and $\check{x}_{t,j,l} $ are the dynamic coefficients of the expansion, constituting the state vector at time $t$. The decomposition enables a separation of the static spatial characteristics, which are considered static functions, that are weighted by dynamic coefficients.

@@

Now we will simplify the notation by defining vectors that contain all the translations of the wavelets and scaling functions 
\begin{align}      
	\boldsymbol\phi_{j_0}(r) \triangleq&\left\lbrace{\phi_{j_0,l}(r)}:l \in \mathbb{Z} \right\rbrace \\
	\boldsymbol\psi_{j}(r) \triangleq& \left\lbrace{\psi_{j,l}(r)}:l \in \mathbb{Z} \right\rbrace\\
	\boldsymbol\alpha_{j_0} \triangleq& \left\lbrace\alpha_{j_0, l}:l \in \mathbb{Z} \right\rbrace \\
	\boldsymbol\beta_{j} \triangleq& \left\lbrace\beta_{j, l}:l \in \mathbb{Z} \right\rbrace \\
	\mathbf{x}_{t,j_0} \triangleq& \left\lbrace{x}_{t,j_{0},l}:l \in \mathbb{Z} \right\rbrace \\
	\check{\mathbf{x}}_{t,j} \triangleq& \left\lbrace{\check{x}}_{t,j,l}:l \in \mathbb{Z} \right\rbrace.
\end{align}
Following this we can define the vectors 
\begin{align}
\boldsymbol\theta^\top \triangleq& [\begin{array}{ccccc} \boldsymbol\alpha_{j_0}^\top & \boldsymbol\beta_{j_0}^\top & \boldsymbol\beta_{j_0+1}^\top & \cdots & \boldsymbol\beta_{J}^\top \end{array}] 
\label{KernelWeights} \\
\mathbf{x}_{t}^\top \triangleq& [\begin{array}{ccccc}\mathbf{x}_{t,j_{0}}^\top &  \check{\mathbf{x}}_{t,j_{0}}^\top & \check{\mathbf{x}}_{t,j_{0}+1}^\top & \cdots & \check{\mathbf{x}}_{t,J}^\top\end{array}]
\label{FieldWeights} \\
\label{KernelBasisVector}
\boldsymbol\lambda^\top(r) \triangleq& \left[
\begin{array}{ccccc} \boldsymbol\phi_{j_0}^\top(r) &
\boldsymbol\psi_{j_0}^\top(r) & 
\boldsymbol\psi_{j_0+1}^\top(r) &
\cdots &
\boldsymbol\psi_{J}^\top(r)\end{array}\right] \\
\label{FieldBasisVector}
\boldsymbol\mu^\top (r) \triangleq& \left[
\begin{array}{ccccc}\boldsymbol\phi_{j_0}^\top(r) &
\boldsymbol\psi_{j_0}^\top(r) & 
\boldsymbol\psi_{j_0+1}^\top(r) &
\cdots &
\boldsymbol\psi_{J}^\top(r)\end{array}\right],
\end{align}
which allow us to write the neural field and connectivity kernel decomposition in the compact form of
\begin{align}
	w\left(r\right) &\approx \boldsymbol\theta^\top\boldsymbol\lambda\left(r\right) 
	\label{eq:KernelFiniteExpansion} \\
	v_t\left(r\right) &\approx \boldsymbol\mu^\top\left(r\right)\mathbf{x}_t.
	\label{eq:FieldFiniteExpansion}
\end{align}

@@

\parham{The state-transition equation can be simplified if the chosen family of wavelets are isotropic where}
\begin{equation}
	\boldsymbol{\lambda}_{j,l}(r-r') = \boldsymbol{\lambda}(2\mathbf{c}_{j,l}+r'-r). 
\end{equation}
Following this, we can switch the order of operations with the inner product and convolution in Eq.~\eqref{eq:ApproxDiscreteTimeModel4}. With this switch the convolution is now between the scaling functions and wavelets for the field and the scaling functions and wavelets for the connectivity kernel. This convolution can be calculated analytically when using B-spline wavelets and can be predefined once, since these are static functions. This has major computation advantages when solving the estimation problem. \parham{Although this switch holds theoretically it should be noted that it will introduce numerical error, specially close to the edges, due to the finite extent of the field. } The predefined convolution can be written as
\begin{equation}\label{eq:Gammaij}
	\left[\Gamma(r')\right]_{:j,l} = T_s \mathbf{\Lambda}_{x}^{-1}\int_{\Omega} \boldsymbol\mu\left(r\right)\boldsymbol\lambda\left(2c_{j,l} + r'-r\right) \,\mathrm{d}r,
\end{equation}
giving
\begin{equation}
	\mathbf{x}_{t+1} = 
	\xi \mathbf{x}_t + 
	\int_{\Omega} \Gamma\left(r'\right)f\left(\boldsymbol\mu^\top\left(r'\right) \mathbf{x}_t\right) 
	\, \mathrm{d}r' \boldsymbol\theta
	+ \mathbf w_t.
\end{equation}

% Substituting Eq.~\eqref{eq:KernelFiniteExpansion} and Eq.~\eqref{eq:FieldFiniteExpansion} in Eq.~\eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(r\right)$, and integrating over the space gives
% \begin{align}\label{eq:DecomposedModel2} 
% 	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1} = 
% 	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
% \end{align}
% Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
% \begin{align}\label{eq:StateEquation}
%  \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
% \label{eq:A_theta}
%  \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I.
% \end{align}
% The disturbance becomes 
% \begin{equation}\label{eq:Disturbance}
% \mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
% \end{equation}
% which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
% \begin{equation}\label{eq:CovMatrix}
% \boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
% \end{equation}

The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into Eq.~\eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix, $\mathbf{C}$, is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ik} \triangleq \int_{\Omega}m(r_i -r')\boldsymbol{\mu}_k(r') \, \mathrm{d}r'.
\end{equation}

\section{B-Spline MRAIDE in State-Space }
\subsection{A Brief Introduction to B-Spline Wavelets}
B-spline functions are one of the most basic elements for wavelet construction. These functions are composed of piecewise polynomials combined smoothly at some breaking points (knots), which are equally spaced in the case of cardinal B-splines. The degree of smoothness is determined by the order of splines \citep{Goswami1999}, where the higher the order the smoother the resulting B-spline becomes (see Fig.~\ref{fig:Figure0}).
\begin{figure}[!t]
\centering
\includegraphics{./Graph/fig3.pdf}
% \includegraphics{./Graph/Figure0b.pdf}
\caption{{\bf Examples of cardinal B-spline functions}. Piecewise polynomial functions (colored lines) are joined at the break points (black dots). (\textbf{A}) B-spline function of order 2 (linear B-spline). (\textbf{B}) B-spline function of order 8.}
\label{fig:Figure0}
\end{figure}

@@

The convolution and inner product of B-spline functions are required in order to formulate the multi-resolution framework described in the previous section (construction of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Gamma$). To show how the convolution of two B-splines is calculated, Eq.~\eqref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ convolutions of the indicator function with itself
\begin{equation}\label{eq:N1convolutions}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
\end{equation}
Using the associativity property of convolution, we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
\end{align}
A direct consequence of Eq.~\eqref{eq:BsplineConvolution} is the inner product between two B-splines is another translated B-spline such that
\begin{align}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
\end{align}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. This holds as the support of $N_m\left(r\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$, i.e. $ N_{m}\left(\frac{m}{2}+r\right)=N_{m}\left(\frac{m}{2}-r\right)$ (see Appendix~\ref{ap:InnerProductOfBsplines} for explicit derivation). The values of $N_{m+m'}$ in Eq.~\eqref{eq:BsplineInnerProduct} can be easily determined recursively by evaluating Eq.~\eqref{eq:MRA-DoBoorFormula} at integer points, i.e.
 \begin{equation}\label{eq:MRA-recursiveBsplineatintegerpoints}
 \begin{cases}
 N_2(k)=\delta(k-1)\quad k\in \mathbb{Z}, \\
 N_{m}\left(k\right)=\frac{k}{m-1}N_{m-1}\left(k\right)+\frac{m-k}{m-1}N_{m-1}\left(k-1\right) \quad k=1,2,\dots,m-1
  \end{cases}
 \end{equation}
Note $N_{m}\left(k\right)=0$ for $k\le0$ or $k\ge m$.

@@

In order to analytically calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$, the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale using the two-scale relation pair defined in equations (\ref{eq:MRA-TwoScalepair1}-\ref{eq:MRA-TwoScalepair2coefs}). For scaling functions this can be done  by usingEq.~\eqref{eq:MRA-TwoScalepair1} recursively, giving
\begin{align}\label{eq:MRA-ConvertingFormulaForscalingFunctions}
 &N_m(r)=\nonumber \\
&\sum_{n_1,n_2, \dots n_j=0}^{m}\left[\mathbf p\right]_{n_1} \left[\mathbf p\right]_{n_2}\dots \left[\mathbf p\right]_{n_j}N_m(2^jr-l_{n_1,n_2, \dots, n_j}),
\end{align}
where 
\begin{align}
 l_{n_1,n_2, \dots, n_j}=2^{j-1}n_1+2^{j-2}n_2+ \dots +2^{0}n_j.
\end{align}
The relation \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} can be also proven by induction (see Appendix~\ref{ap:Induction}). A similar method can be used for wavelet functions noting that they must be written first in terms of scaling functions using Eq.~\eqref{eq:MRA-TwoScalepair2}. Examples of such expansions are given in Fig.~\ref{fig:MRA-BasisDecomposition} where the scaling and wavelet functions at $j=0$ are written in terms of 625 and 1250 scaling functions at $j=4$ respectively.
\begin{figure}[!t] 
 	\centering
 		\includegraphics[scale=1]{./Graph/fig5.pdf}
 		\caption{{\bf The scaling and the wavelet function decomposition}. (\textbf{A}) The scaling function (blue) at $j=0$ is obtained using 625 scaling functions (black) at $j=4$. (\textbf{B}) The wavelet function (blue) at $j=0$ is obtained using 1250 scaling functions (black) at $j=4$.}
 	\label{fig:MRA-BasisDecomposition}
 \end{figure}  
 In this paper, 4th -- order cardinal B-spline scaling and wavelet functions are used, resulting in  8th and 12th order B-spline functions when calculating Eq.~\eqref{eq:Lambdax} and Eq.~\eqref{eq:Uij} respectively  whose values at integer points are given in Table~\ref{table:MRA-BsplineatIntegerPoints}.
\begin {table}[t]
\begin{center}
	\begin{tabular}{lcccc}
	\hline \hline
	& $k$ & $3!N_{4}\left(k\right)$ & $7!N_{8}\left(k\right)$ & $11!N_{12}\left(k\right)$\\ 
	\hline 
	& 1 & $1$ & $1$ & $1$\\
	& 2 & $4$ & $120$ & $2,036$\\
	& 3 & $\space$ & $1,191$ & $152,637$\\
	& 4 & $\space$ & $2,416$ & $2,203,488$\\
	& 5 & $\space$ & $\space$ & $9,738,114$\\
	& 6 & $\space$ & $\space$ & $15,724,248$\\
	\hline \hline
	\end{tabular}
 \caption {{\bf Cardinal B-Splines at the knot sequence}. Table entries are calculated using Eq.~\eqref{eq:MRA-recursiveBsplineatintegerpoints} and can be also found in \citet{Goswami1999}.} 
 \label{table:MRA-BsplineatIntegerPoints}
 \end{center}
 \end {table}

>>>>

The multi-resolution approximation \citep{Mallat1989a} of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(\cdot)$, and the connectivity kernel, $w(\cdot)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(\mathbf{r})$ and a mother wavelet $\psi(\mathbf{r})$. Considering a  one-dimensional field, the static connectivity kernel is decomposed as,
\begin{equation}
 w\left(r\right)=\sum_{l\in \mathbb{Z}}\alpha_{j_0,l} \underbrace{\phi_{j_0,l}\left(r\right)}_{\text{scaling functions}} + \sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}}\beta_{j,l} \underbrace{\psi_{j,l}\left(r\right)}_{\text{wavelets}}, 
\label{eq:KernelExpansion}
\end{equation}
where $\alpha_{j_0,l}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,l}$ are the detail coefficients at different scales $j$, with
\begin{align}
\phi_{j,l}\left(r\right) =& 2^{\frac{j}{2}}\phi\left(2^j r-l\right) \\	
\psi_{j,l}\left(r\right) =& 2^{\frac{j}{2}}\psi\left(2^j r-l\right).
\end{align}
The parameters $j$ and $l$ control the scale (dilation) and translation (spatial shift), respectively. \dean{NEED SOMETHING ABOUT HOW j AND l ARE SET HERE - AS IN THAT THEY ARE KNOWN - AND THAT WE ONLY NEED TO ESTIMATE THE COEFFICIENTS.} \parham{The number of spatial shifts is dependent on the scale, $j$, and can be chosen such that the basis functions cover the spatial domain of interest. The number of scales depends on the spectral characteristics of the field, which will be described later.} Scaling functions retain the lowest frequency components, due to their low-pass properties, while the wavelet functions extract successively higher frequency components due to their band-pass characteristics. 
% At each spatial scale indexed by $j$ (level of decomposition), the neural field can be written as
% \begin{equation}
% 	v_{j,t}\left(\mathbf{r}\right) = \begin{cases}
% 	\sum_{l \in \mathbb{Z}}x_{j,l}(t) \underbrace{\phi_{j,l}\left(\mathbf{r}\right)}_{\text{scaling function}} + \check{x}_{t,j,l}\underbrace{\psi_{j,l}\left(\mathbf{r}\right)}_{\text{mother wavelet}} & \text{if $j=0$},\\
% 	\sum_{l \in \mathbb{Z}}\check{x}_{t,j,l}\psi_{j,l}\left(\mathbf{r}\right) & \text{otherwise}
% \end{cases}
% \end{equation}
The dynamic neural field is decomposed using
\begin{equation}
 v_t\left(r\right)=\sum_{l \in \mathbb{Z}}x_{t,j_{0},l} \underbrace{\phi_{j_0,l}\left(r\right)}_{\text{scaling functions}} + \sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}^n} \check{x}_{t,j,l} \underbrace{\psi_{j,l}\left(r\right)}_{\text{wavelets}},
\label{eq:FieldExpansion}
\end{equation}
where $x_{t,j_{0},l}$ and $\check{x}_{t,j,l} $ are the dynamic coefficients of the expansion, constituting the state vector at time $t$. The decomposition enables a separation of the static spatial characteristics, which are considered static functions, that are weighted by dynamic coefficients.

@@

Now we will simplify the notation by defining vectors that contain all the translations of the wavelets and scaling functions 
	\boldsymbol\phi_{j_0}(r) \triangleq&\left\lbrace{\phi_{j_0,l}(r)}:l \in \mathbb{Z} \right\rbrace \\
	\boldsymbol\psi_{j}(r) \triangleq& \left\lbrace{\psi_{j,l}(r)}:l \in \mathbb{Z} \right\rbrace\\
	\boldsymbol\alpha_{j_0} \triangleq& \left\lbrace\alpha_{j_0, l}:l \in \mathbb{Z} \right\rbrace \\
	\boldsymbol\beta_{j} \triangleq& \left\lbrace\beta_{j, l}:l \in \mathbb{Z} \right\rbrace \\
	\mathbf{x}_{t,j_0} \triangleq& \left\lbrace{x}_{t,j_{0},l}:l \in \mathbb{Z} \right\rbrace \\
	\check{\mathbf{x}}_{t,j} \triangleq& \left\lbrace{\check{x}}_{t,j,l}:l \in \mathbb{Z} \right\rbrace.
Following this we can define the vectors 
\boldsymbol\theta^\top \triangleq& [\begin{array}{ccccc} \boldsymbol\alpha_{j_0}^\top & \boldsymbol\beta_{j_0}^\top & \boldsymbol\beta_{j_0+1}^\top & \cdots & \boldsymbol\beta_{J}^\top \end{array}] 
\label{KernelWeights} \\
\mathbf{x}_{t}^\top \triangleq& [\begin{array}{ccccc}\mathbf{x}_{t,j_{0}}^\top &  \check{\mathbf{x}}_{t,j_{0}}^\top & \check{\mathbf{x}}_{t,j_{0}+1}^\top & \cdots & \check{\mathbf{x}}_{t,J}^\top\end{array}]
\label{FieldWeights} \\
\label{KernelBasisVector}
\boldsymbol\lambda^\top(r) \triangleq& \left[
\begin{array}{ccccc} \boldsymbol\phi_{j_0}^\top(r) &
\boldsymbol\psi_{j_0}^\top(r) & 
\boldsymbol\psi_{j_0+1}^\top(r) &
\cdots &
\boldsymbol\psi_{J}^\top(r)\end{array}\right] \\
\label{FieldBasisVector}
\boldsymbol\mu^\top (r) \triangleq& \left[
\begin{array}{ccccc}\boldsymbol\phi_{j_0}^\top(r) &
\boldsymbol\psi_{j_0}^\top(r) & 
\boldsymbol\psi_{j_0+1}^\top(r) &
\cdots &
\boldsymbol\psi_{J}^\top(r)\end{array}\right],
which allow us to write the neural field and connectivity kernel decomposition in the compact form of
	w\left(r\right) &\approx \boldsymbol\theta^\top\boldsymbol\lambda\left(r\right) 
	\label{eq:KernelFiniteExpansion} \\
	v_t\left(r\right) &\approx \boldsymbol\mu^\top\left(r\right)\mathbf{x}_t.
	\label{eq:FieldFiniteExpansion}

@@

\parham{The state-transition equation can be simplified if the chosen family of wavelets are isotropic where}
	\boldsymbol{\lambda}_{j,l}(r-r') = \boldsymbol{\lambda}(2\mathbf{c}_{j,l}+r'-r). 
Following this, we can switch the order of operations with the inner product and convolution in Eq.~\eqref{eq:ApproxDiscreteTimeModel4}. With this switch the convolution is now between the scaling functions and wavelets for the field and the scaling functions and wavelets for the connectivity kernel. This convolution can be calculated analytically when using B-spline wavelets and can be predefined once, since these are static functions. This has major computation advantages when solving the estimation problem. \parham{Although this switch holds theoretically it should be noted that it will introduce numerical error, specially close to the edges, due to the finite extent of the field. } The predefined convolution can be written as
\begin{equation}\label{eq:Gammaij}
	\left[\Gamma(r')\right]_{:j,l} = T_s \mathbf{\Lambda}_{x}^{-1}\int_{\Omega} \boldsymbol\mu\left(r\right)\boldsymbol\lambda\left(2c_{j,l} + r'-r\right) \,\mathrm{d}r,
giving
	\int_{\Omega} \Gamma\left(r'\right)f\left(\boldsymbol\mu^\top\left(r'\right) \mathbf{x}_t\right) 
	\, \mathrm{d}r' \boldsymbol\theta
	+ \mathbf w_t.
% Substituting Eq.~\eqref{eq:KernelFiniteExpansion} and Eq.~\eqref{eq:FieldFiniteExpansion} in Eq.~\eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(r\right)$, and integrating over the space gives
% \begin{align}\label{eq:DecomposedModel2} 
% 	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1} = 
% 	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
% Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
% \begin{align}\label{eq:StateEquation}
%  \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
% \label{eq:A_theta}
%  \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I.
% The disturbance becomes 
% \begin{equation}\label{eq:Disturbance}
% \mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
% \end{equation}
% which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
% \begin{equation}\label{eq:CovMatrix}
% \boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
% \end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into Eq.~\eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
where each element of the observation matrix, $\mathbf{C}$, is given by
\begin{equation}\label{eq:Observationmatrix}
	\mathbf{C}_{ik} \triangleq \int_{\Omega}m(r_i -r')\boldsymbol{\mu}_k(r') \, \mathrm{d}r'.
\section{B-Spline MRAIDE in State-Space }
\subsection{A Brief Introduction to B-Spline Wavelets}
B-spline functions are one of the most basic elements for wavelet construction. These functions are composed of piecewise polynomials combined smoothly at some breaking points (knots), which are equally spaced in the case of cardinal B-splines. The degree of smoothness is determined by the order of splines \citep{Goswami1999}, where the higher the order the smoother the resulting B-spline becomes (see Fig.~\ref{fig:Figure0}).
\includegraphics{./Graph/fig3.pdf}
% \includegraphics{./Graph/Figure0b.pdf}
\caption{{\bf Examples of cardinal B-spline functions}. Piecewise polynomial functions (colored lines) are joined at the break points (black dots). (\textbf{A}) B-spline function of order 2 (linear B-spline). (\textbf{B}) B-spline function of order 8.}
\label{fig:Figure0}

@@

The convolution and inner product of B-spline functions are required in order to formulate the multi-resolution framework described in the previous section (construction of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Gamma$). To show how the convolution of two B-splines is calculated, Eq.~\eqref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ convolutions of the indicator function with itself
\begin{equation}\label{eq:N1convolutions}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
\end{equation}
Using the associativity property of convolution, we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
\end{align}
A direct consequence of Eq.~\eqref{eq:BsplineConvolution} is the inner product between two B-splines is another translated B-spline such that
\begin{align}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
\end{align}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. This holds as the support of $N_m\left(r\right)$ is $\left[ 0,m\right]$ and  is symmetric with respect to $r=\frac{m}{2}$, i.e. $ N_{m}\left(\frac{m}{2}+r\right)=N_{m}\left(\frac{m}{2}-r\right)$ (see \ref{ap:InnerProductOfBsplines} for explicit derivation). The values of $N_{m+m'}$ in Eq.~\eqref{eq:BsplineInnerProduct} can be easily determined recursively by evaluating Eq.~\eqref{eq:MRA-DoBoorFormula} at integer points, i.e.
 \begin{equation}\label{eq:MRA-recursiveBsplineatintegerpoints}
 \begin{cases}
 N_2(k)=\delta(k-1)\quad k\in \mathbb{Z}, \\
 N_{m}\left(k\right)=\frac{k}{m-1}N_{m-1}\left(k\right)+\frac{m-k}{m-1}N_{m-1}\left(k-1\right) \quad k=1,2,\dots,m-1
  \end{cases}
 \end{equation}
Note $N_{m}\left(k\right)=0$ for $k\le0$ or $k\ge m$.

@@

In order to analytically calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$, the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale using the two-scale relation pair defined in equations (\ref{eq:MRA-TwoScalepair1}-\ref{eq:MRA-TwoScalepair2coefs}). For scaling functions this can be done  by usingEq.~\eqref{eq:MRA-TwoScalepair1} recursively, giving
\begin{align}\label{eq:MRA-ConvertingFormulaForscalingFunctions}
 &N_m(r)=\nonumber \\
&\sum_{n_1,n_2, \dots n_j=0}^{m}\left[\mathbf p\right]_{n_1} \left[\mathbf p\right]_{n_2}\dots \left[\mathbf p\right]_{n_j}N_m(2^jr-l_{n_1,n_2, \dots, n_j}),
 l_{n_1,n_2, \dots, n_j}=2^{j-1}n_1+2^{j-2}n_2+ \dots +2^{0}n_j.
The relation \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} can be also proven by induction (see \ref{ap:Induction}). A similar method can be used for wavelet functions noting that they must be written first in terms of scaling functions using Eq.~\eqref{eq:MRA-TwoScalepair2}. Examples of such expansions are given in Fig.~\ref{fig:MRA-BasisDecomposition} where the scaling and wavelet functions at $j=0$ are written in terms of 625 and 1250 scaling functions at $j=4$ respectively.
 		\includegraphics[scale=1]{./Graph/fig5.pdf}
 		\caption{{\bf The scaling and the wavelet function decomposition}. (\textbf{A}) The scaling function (blue) at $j=0$ is obtained using 625 scaling functions (black) at $j=4$. (\textbf{B}) The wavelet function (blue) at $j=0$ is obtained using 1250 scaling functions (black) at $j=4$.}
 	\label{fig:MRA-BasisDecomposition}
 In this paper, 4th -- order cardinal B-spline scaling and wavelet functions are used, resulting in  8th and 12th order B-spline functions when calculating Eq.~\eqref{eq:Lambdax} and Eq.~\eqref{eq:Uij} respectively  whose values at integer points are given in Table~\ref{table:MRA-BsplineatIntegerPoints}.
\begin {table}[t]
\begin{center}
	\begin{tabular}{lcccc}
	\hline \hline
	& $k$ & $3!N_{4}\left(k\right)$ & $7!N_{8}\left(k\right)$ & $11!N_{12}\left(k\right)$\\ 
	& 1 & $1$ & $1$ & $1$\\
	& 2 & $4$ & $120$ & $2,036$\\
	& 3 & $\space$ & $1,191$ & $152,637$\\
	& 4 & $\space$ & $2,416$ & $2,203,488$\\
	& 5 & $\space$ & $\space$ & $9,738,114$\\
	& 6 & $\space$ & $\space$ & $15,724,248$\\
	\hline \hline
 \caption {{\bf Cardinal B-Splines at the knot sequence}. Table entries are calculated using Eq.~\eqref{eq:MRA-recursiveBsplineatintegerpoints} and can be also found in \citet{Goswami1999}.} 
 \label{table:MRA-BsplineatIntegerPoints}
 \end{center}
 \end {table}
\subsection{Frequency Analysis}
To determine the level of approximation in the model, $j$, detailed in the previous section, that is the number of basis functions required to reconstruct the field, $v_t(r)$, the frequency response of the B-spline function needs to be computed by applying the Fourier transform to Eq.~\eqref{eq:N1convolutions}. To proceed the Fourier transform of $N_1(r)$ is first calculated, 
\begin{align}\label{eq:MRA-N1Fouriertransform}
\mathcal F(N_1(r))&=\int_{-\infty}^{+\infty}N_1(r)\mathrm{exp}(-2\pi i \nu r)dr \nonumber \\
&=\int_{0}^{1} \mathrm{exp}(-2\pi i \nu r)dr \nonumber \\
&=\frac{1-\mathrm{exp}(-2\pi i \nu)}{2\pi i\nu}.
